### 2013-12-11 supplemental.tex

\documentclass[article]{llncs}

% preprint only
\pagestyle{plain}
\usepackage{amsmath,stmaryrd,listings,mathpartir,amssymb,graphicx}

\usepackage[usenames,dvipsnames]{color}

\usepackage{pfsteps} % for appendix

\newcommand{\Scribtexttt}[1]{{\texttt{#1}}}
\newcommand{\SColorize}[2]{\color{#1}{#2}}
\newcommand{\inColor}[2]{{\Scribtexttt{\SColorize{#1}{#2}}}}
\definecolor{PaleBlue}{rgb}{0.90,0.90,1.0}
\newcommand{\rackett}[1]{\inColor{black}{#1}}
\input{local-macros}
\usepackage[numbers,sectionbib]{natbib}
\usepackage[english]{babel}
% pdfpagescrop for preprint only
%\usepackage[hidelinks,pdfpagescrop={92 62 523 728}]{hyperref}
\usepackage[hidelinks]{hyperref}
\usepackage{cleveref}
\usepackage{placeins,xr}

\externaldocument{paper-jfp}

\input{preamble}
\begin{document}
\title{The Essence of Summarization\\ in Pushdown Flow Analyses\\
       \textbf{Supplemental Material}}
\author{J. Ian Johnson and David Van Horn}

\section{PDCFA}
\input{metatheory}

\begin{theorem}\label{thm:concrete-sr}
  Tabular semantics simulates the standard semantics given fresh allocation.
\end{theorem}
\begin{proof}[Sketch]
  We have the invariant that the continuation and meta-continuation have unique unrollings in addition to the table invariants from paper's section 3,
  since the distinct $\mstore$ at push and capture time makes entries in $\mktab$ unique, regardless of $\mmktab$.
\end{proof}

\noindent{}Proof of paper's theorem ``Correctness of PDCFA''
\begin{proof}[Sketch]
  The invariants on the tables still hold. 
  The widening to share $\mmemo$ and $\mktab$ amongst all explored states reifies the search state into a $\sa{System}$. We can add a set of edges simply.
  If we add $\mmemo$ and $\mktab$ to all states in these edges, we get a reduction relation comparable to the unwidened tabular semantics.
  We prove that the infinite unrolling of the standard reduction relation stuttering bisimulates the relation reified from an infinite stepping of the widened tabular semantics.
  This only holds if the state selection from the frontier is fair, since unfair choices can leave some function calls unresolved.
  Formally,
  \begin{equation*}
    \setbuild{\mtrace}{\mtrace : i \le \omega, \mstate_0 \longmapsto^i \mstate} \simeq {\mathcal F}(\mstate_0)^\omega
  \end{equation*}
  where $\simeq$ is TODO
by coinduction and appeals to the table invariants.
  If a new function result is found after a call already used the memo table, the context of the memo user is used after memoizing, so it looks like that result was part of the original memo table lookup.
  The fair selection guarantees that all function results will be eventually discovered, so the memo table appeals eventually catch up to the found results.
\end{proof}

\begin{theorem}[No dangling pointers]\label{thm:gc-pointers}
  All addresses in $\Gamma(\mstate)$ are in the domain of $\Gamma(\mstate).\mstore$.
\end{theorem}

\begin{theorem}[Garbage is irrelevant]\label{thm:gc-concrete}
  If $\mstate \stepto^n \mstate'$ and $\mstate \stepto^n \mstate''$ then $\Gamma(\mstate') = \Gamma(\mstate'')$.
\end{theorem}

\begin{theorem}[Abstract GC is sound for garbage-free states]\label{thm:gc-sound}
  If $\mstate \stepto \mstate'$ and $\alpha(\Gamma(\mstate)) \sqsubseteq \mastate$ then there is a $\mastate'$ such that $\mastate \astepto \mastate'$ and $\alpha(\Gamma(\mstate')) \sqsubseteq \mastate'$.
\end{theorem}

This final proposition is not in the original work on abstract GC; they prove soundness with respect to a concrete semantics that performs GC after every step.
%
The moral correctness of the abstraction is that it never under-approximates \emph{relevant} portions of the state space --- in this case, the reachable subset of the store.
%
The definitions of $\alpha$, $\astepto$ and $\sqsubseteq$ are the simple pointwise abstractions of addresses through states and the reduction relation, and finally a lifted subset relation between states and their components, respectively.
%
Both $\alpha$ and $\sqsubseteq$ are parameterized on the choice of $\Addr$, for which the following soundness criterion should hold:
\begin{mathpar}
  \inferrule*[right={[Sound allocation]}]{\alpha(\mstate) \sqsubseteq \mastate}{\alpha(\alloc(\mstate)) \sqsubseteq \widehat{\alloc}(\mastate)}
\end{mathpar}

\section{Shift/reset}
\noindent{}Proof of paper's theorem 4 follows the same line of reasoning as the correctness of PDCFA.

% \noindent{}Proof of \autoref{thm:pdcfa-tabular}, which is more formally stated as (tabular $\pdstepto$, CESK $\stepto$): There is a well-founded equivalence bisimulation between $\pdstepto$ (quotiented to states satisfying $\inv$) and $\tred{\stepto}$.
% We define equivalence ($\simeq$) as
% \begin{mathpar}
%   \inferrule{\mkont \in \unroll{\mktab}{\mkont'} \\ \mkont \nequiv \kcons{\kfn{\mval}}{\mkont''}}
%             {\mtrace\tpl{\mpoint,\mstore,\mkont} \simeq \tpl{\mpoint,\mstore,\mkont,\mktab,\mmemo}} \\
% %
%   \inferrule{\mkont \in \unroll{\mktab}{\mkont'} \\
%              ((\mexpr,\menv[\mvar \mapsto \maddr]), \mstore\sqcup[\maddr \mapsto \set{\mval}]) \notin \dom(\mmemo)}
%             {\mtrace\tpl{\mval,\mstore,\kcons{\kfn{\slam{\mvar}{\mexpr}, \menv}{\mkont}}} \simeq \tpl{\mval,\mstore,\kcons{\kfn{\slam{\mvar}{\mexpr},\menv}}{\mkont'},\mktab,\mmemo}} \\
% %
%   \inferrule{\tpl{\mval', \mstore', \mkont} \notin \mtrace'' \\
%              \mkont \in \unroll{\mktab}{\mkont''} \\
%              ((\mexpr,\menv[\mvar \mapsto \maddr]), \mstore\sqcup[\maddr \mapsto \set{\mval}]) \in \dom(\mmemo)}
%             {\mtrace'\tpl{\mval, \mstore, \kcons{\kfn{\slam{\mvar}{\mexpr}, \menv}}{\mkont}}\mtrace''
%               \simeq \tpl{\mval, \mstore, \kcons{\kfn{\slam{\mvar}{\mexpr}, \menv}}{\mkont''}, \mktab, \mmemo}}
% \end{mathpar}
% This means that states which ``line up'' regardless of history are related, and that states that follow a function call that has not yet completed are related to the function call state where we already have the memoized results.
% %
% Each step in this case will get ``closer to'' the result to memoize, as given by the length of the path we get from the invariant on the memo table.
% \begin{proof}
%   Let $\mtrace,\mtrace',\mstate_\Xi$ be arbitrary such that $\mtrace \simeq \mstate_\Xi$ and $\mtrace \tred{\stepto} \mtrace'$.
%   By cases on $\mtrace \simeq \mstate_\Xi$:
%   \begin{byCases}
%     \case{\text{First rule}}{One step related by simple case analysis.}
%     \case{\text{Second rule}}{One step related by definitions of $\mathit{unroll}$, $\stepto$ and $\pdstepto$.}
%     \case{\text{Third rule}}{
%     By cases on $\mtrace\mstate \tred{\stepto} \mtrace\mstate\mstate'$:
%     \begin{byCases}
%       \case{\mtrace\tpl{(\svar\mvar, \menv), \mstore, \mkont} \stepto
%         \mtrace\mstate\tpl{\mval,\mstore,\mkont}}{...}
%       % application
%       \case{\mtrace\tpl{(\sapp{\mexpri0}{\mexpri1}, \menv), \mstore,
%           \mkont} \stepto \mtrace\mstate\tpl{(\mexpri0, \menv),
%           \mstore, \kcons{\kar{\mexpri1,\menv}}{\mkont}}}{...}
%       % argument evaluation
%       \case{\mtrace\tpl{\mval, \mstore,
%           \kcons{\kar{\mexpr,\menv}}{\mkont}} \stepto
%         \mtrace\mstate\tpl{(\mexpr, \menv), \mstore,
%           \kcons{\kfn{\mval}}{\mkont}}}{...}
%       % function call
%       \case{\mtrace\tpl{\mval,\mstore,\kcons{\kfn{\vclo{\slam{\mvar}{\mexpr}}{\menv}}}{\mkont}}
%         \stepto \mtrace\mstate\tpl{(\mexpr,
%           \extm{\menv}{\mvar}{\maddr}),
%           \joinone{\mstore}{\maddr}{\mval}, \mkont}}{...}
%     \end{byCases}}
%   \end{byCases}
%   By \autoref{lem:tab-inv}

%   Soundness is a corollary of \autoref{lem:tab-inv}.
% \end{proof}

% \noindent{}Proof of \autoref{thm:global}.
% \begin{proof}
% %% TODO
%   The invariants of the memo and k-tables are independent of the states that carry them, so lifting them to a global position does not affect their information.
% \end{proof}

% \noindent{}Proof of \autoref{thm:refinement}.
% \begin{proof}
%   todo
% \end{proof}

% \noindent{}Proof of \autoref{thm:cfa2}.
% \begin{proof}
%   todo
% \end{proof}

% \noindent{}Proof of \autoref{thm:concrete-sr}.
% \begin{proof}
%   todo
% \end{proof}

% \noindent{}Proof of \autoref{thm:sound-sr}
% \begin{proof}
%   todo
% \end{proof}

\end{document}